High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo

### 2022-05-18 PNAS_manuscript/PNAS_manuscript.tex
edits for final files

\significancestatement{Acquiring high-fidelity ancient magnetic field intensity records from rocks is crucial for constraining the long-term evolution of Earth's core. However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal behavior. We use anorthosites that formed in the deep crust and were brought to the near surface where they acquired thermal remanent magnetizations. These anorthosites have experienced minimal post-formation alteration and yield high-quality paleointensity estimates. In contrast to scenarios of a progressively decaying field leading up to late nucleation of Earth's inner core, these data record a strong field 1.1 Ga. A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo during the late Mesoproterozoic Era.}

>>>>

\significancestatement{Acquiring high-fidelity ancient magnetic field intensity records from rocks is crucial for constraining the long-term evolution of Earth's core. However, robust estimates of ancient field strengths are often difficult to recover due to alteration or non-ideal behavior. We use rocks known as anorthosite that formed in the deep crust and were brought to the near surface where they acquired thermal remanent magnetizations. These rocks have experienced minimal post-formation alteration and yield high-quality paleointensity estimates. In contrast to scenarios of a progressively decaying field leading up to proposed late nucleation of Earth's inner core, these data record a strong field 1.1 billion years ago. A strong field that persisted over a 14-million-year interval indicates the existence of appreciable power sources for Earth's dynamo at this time.}
